### 2023-04-15 paper/language.tex
Translation into PikaCore

\section{Syntax}

\subsection{Shared Syntax}

The following syntactic forms are shared between the \Pika{} surface language and \PikaCore.

\begin{align*}
  &i ::= \cdots \mid -2 \mid -1 \mid 0 \mid 1 \mid 2 \mid \cdots
  \\
  &b ::= \True{} \mid \False
  \\
  &e ::= x \mid i \mid b \mid \lambda x.\; e
  \\
  &\tau ::= \Int \mid \Bool \mid \tau \ra \tau
\end{align*}

\subsection{\Pika{} Syntax}
\begin{align*}
  &e ::= \cdots \mid \apply_{\alpha}(e) \mid \liftExpr_{\alpha}(e) \mid \Lambda (\alpha \isA \layout{X}).\; e
  \\
  &\tau ::= \cdots \mid X \mid \alpha \mid e\; e \mid (\alpha \isA \layout{X}) \Ra \tau
\end{align*}

\subsection{\PikaCore{} Syntax}
\begin{align*}
  &e ::= \cdots \mid e\; \overline{a} \mid \sslmath{\overline{x}}{\overline{h}} \mid \withIn{\{ \overline{x} \} := e} e
  \\
  &a ::= \{ \overline{x} \} \mid \{ i \} \mid \{ b \}
  \\
  &\tau ::= \cdots \mid \SSL(n)
  \\
  &h ::= \ell \mapsto e
  \\
  &\ell ::= x \mid (x + n)
  \\
  &n ::= 0 \mid 1 \mid 2 \mid \cdots
\end{align*}

The construct $\sslmath{\overline{x}}{\overline{h}}$ is an SSL assertion with heaplets $\overline{h}$ and
bound variables $\overline{x}$. Note that it can have free variables as well.

% The general form of a one-parameter \PikaCore{} function definition is
% \begin{align*}
%   &f : \SSL(m) \ra \SSL(n)\\
%   &f\; \{ \overline{\ell \mapsto y} \} \outs \{ \overline{r} \} := e\\
%   &f\; \{ \overline{\ell' \mapsto y} \} \outs \{ \overline{r} \} := e'\\
%   &\cdots
% \end{align*}
%
% \noindent
% where $\overline{r}$ are the names of the output parameters. These output parameters
% must be the same in each branch and it is also required that the number of output parameters matches the type signature: $|\overline{r}| = n$.
%
% Output parameters are given in \PikaCore{} as part of the \verb|with-in| construct:
%
% \begin{align*}
%   \withIn{\{ a, b \} := f\; \{ x \}} e
% \end{align*}
%
% This applies the function $f$ to the input parameter $x$ and gives $a$ and $b$ as output parameters. These two
% output parameters of the application are brought into scope for the \PikaCore{} expression $e$.
%