### 2023-08-21 new-paper/examples.tex

\section{Examples}
\label{sec:examples}

\subsection{\Pika{} 1}

\subsection{\Pika{} 2}

Version 2 of \Pika{} is a rewrite of version 1 and it is missing a couple of features of version 1. However, it has the \synth{} feature. It also has the groundwork for layout polymorphism to be implemented. The second aspect, in particular, is a substantial change from version 1 and motivated the rewrite.