### 2023-08-21 new-paper/semantics.tex

\section{Formal Semantics}
\label{sec:semantics}