### 2023-09-06 new-paper/soundness.tex
soundness

\subsection{From \tool to SLL Specifications: Soundness of the Translation}
\label{sec:soundness}

We want to show that our abstract machine semantics and our SSL
translation fit together. In particular, our abstract machine
semantics should generate models that satisfy the separation logic
propositions given by our SSL translation.
Figure~\ref{fig:soundness-diagram} gives a high-level overview of how
these pieces fit together. We will give a more specific description of
this in Theorem~\ref{thm:gen-soundness}.

\input{soundness-diagram.tex}

\begin{restatable}[Soundness]{theorem}{soundnessThm} \label{thm:gen-soundness}
  For any well-typed expression $e$, the following holds where $\VarSet = \dom(\EndSigma)$:
  \begin{itemize}
    \item If $\Tsem{e}{\VarSet, r}$ is satisfiable and $(e, \StartSigma, \Start{h}, \mathcal{F}) \step (e', \EndSigma, \End{h'}, \mathcal{F}', r)$\\
          then $\EndPair \models \Tsem{e}{\VarSet, r}$
  \end{itemize}
  That is, given an expression $e$ with a satisfiable SSL translation, any heap machine state that $e$ transitions to (by the abstract machine semantics) will
  be a model for the SSL translation of $e$. See Figure~\ref{fig:soundness-diagram}.
\end{restatable}

\begin{proof}
  See the Appendices
  %~\ref{sec:soundness-proof}
  in the extended version of the paper.
\end{proof}

The fact that, at the top level, we only translate function definitions suggests an additional theorem. We want to specifically
show that any possible function application is sound, in the sense just described. This immediately follows from Theorem~\ref{thm:gen-soundness}.

Abbreviating $\instantiateS{A,B}{f}$ as $f_{A,B}$:

\begin{theorem}[Application soundness] \label{thm:fn-soundness}
  For any well-typed function application $f_{A,B}(e)$, the following should hold where $V = \dom(\EndSigma)$:
  \begin{itemize}
    \item If $\Tsem{f_{A,B}(e)}{\VarSet, r}$ is satisfiable and $(f_{A,B}(e), \StartSigma, \Start{h}, \mathcal{F}) \step (e', \EndSigma, \End{h'}, \mathcal{F}', r)$\\
          then $\EndPair \models \Tsem{f_{A,B}(e)}{\VarSet, r}$
  \end{itemize}
\end{theorem}

\begin{proof}
  This follows immediately from Theorem~\ref{thm:gen-soundness}.
\end{proof}