### 2023-10-03 writing/new-paper/conclusion.tex
Tidy up to be able to navigate
src is pika source (we ignore c bit)
suslik is the underlying suslik (may need updates)
tests has all the tests we are trying to get to work
writing is the paper we want to submit/revise
various is odd bits and pieces and old C stuff

\section{Conclusion}
\label{sec:conclusion}

We have seen a technique for generating an SSL specification from a functional program, such that this translation is sound with respect to the standard semantics. In order to do this, we built the notion of a \textit{layout} of an algebraic data type on top of separation logic. Then we observed that \SuSLik, the tool we use to synthesise programs from SSL, has notable limitations in its ability to synthesise the SSL specifications we generate. This motivated the creation of the \synth{} keyword. This new keyword allows us to have direct access to the synthesis features of \SuSLik{} from inside the \Pika{} programming language.