### 2023-10-03 writing/new-paper/future-work.tex

\section{Future Work}
\label{sec:future-work}

One direction for future work is the extension of \Pika{} 2 to include support for higher-order functions and finishing the implementation of layout polymorphic type checking.

The reverse transformation deserves further investigation: if we go from
an SSL specification to \Pika{} program and then compile to, \eg, C, can
we synthesise additional programs that a traditional SSL synthesisers
would struggle with? What are the limitations of this approach?

Finally, is it possible to derive translations for languages such as \Pika{}
from abstract machine semantics? In this paper, we have given a
language with abstract machine semantics. We then give a translation
of that language into SSL. We then show that the final states given by
the abstract machine semantics are models for the SSL propositions
produced by our translation. But is it possible to begin by specifying the
abstract machine semantics and then mathematically (or automatically)
\textit{derive} an appropriate translation into SSL, with the
requirement that the translation satisfies the soundness theorem?