### 2023-10-03 writing/notes-on-pikac/introduction.tex

\section{Introduction}

\Pika{} is a functional programming language with layouts and layout
polymorphism. \PikaCore{} is an intermediate representation of \Pika{} where all
layout polymorphism has been removed. That is, in \PikaCore, all layout
polymorphism is specialized to specific layouts and SSL assertions are
first-class values. Furthermore, \PikaCore{} is in ANF form.

From \PikaCore, the program can be translated to several different targets. This
pipeline is given in Fig.~\ref{fig:translation-pipeline}.

\begin{figure}
  \input{pipeline.tex}
  \caption{Translation pipeline}
  \label{fig:translation-pipeline}
\end{figure}